---
title: "Reproducible Peer Assessment 1"
output: html_document
---
##Loading and preprocessing the data
```{r}
#Load the data
setwd('C:/Users/Zhongyu/Documents/learning/Rep Proj 1');
data <- read.csv('activity.csv')
```
##What is mean total number of steps taken per day?
```{r}
#Calculate the total number of steps taken per day
steps <- aggregate(data$steps, by=list(date=data$date), FUN=sum)
#Make a histogram of the total number of steps taken each day
hist(steps$x, main='Histogram of Steps per Day')
#Calculate and report the mean and median of the total number of steps taken per day
mean_steps <- as.integer(mean(steps$x, na.rm=TRUE))
median_steps <- as.integer(median(steps$x, na.rm=TRUE))
```
The mean and median value of total number of steps taken per day is `r mean_steps` and `r median_steps`


##What is the average daily activity pattern?
```{r}
#Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
meanrm <- function (x) {
  mean(x,na.rm=TRUE)
}
steps_by_hour <- aggregate(data$steps, by=list(hour=data$interval), FUN=meanrm)
steps_by_hour$x[is.na(steps_by_hour$x)] = 0
plot(steps_by_hour$x, type = 'l', main= 'Average Steps per 5 Minute Range')
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
index <- which.max(steps_by_hour$x)
time <- steps_by_hour$hour[index]
value <- steps_by_hour$x[index]
```
The max step is `r value` at time `r time`


##Imputing missing values
The method I use for inputation is simply replace all missing values to zero.
It is reasonable to do so because it is probaby because that there are no steps at all causes value missing. Because human motion can change a lot over time it is hard to find a good imputation method.
```{r}
#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
number_na = sum(is.na(data$steps))
#Devise a strategy for filling in all of the missing values in the dataset.
#Create a new dataset that is equal to the original dataset but with the missing data filled in.
data_rm <- data;
data_rm$steps[is.na(data$steps)] = 0
#Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
steps_rm <- aggregate(data_rm$steps, by=list(data_rm$date), FUN=sum)
hist(steps_rm$x, main='Histogram of Steps per Day')
mean_steps <- mean(steps_rm$x, na.rm=TRUE)
median_steps <- median(steps_rm$x, na.rm=TRUE)

```
The mean and median value of total number of steps taken per day is `r mean_steps` and `r median_steps`. It is lower than expected because all missing values are assigned 0, which is definitely lower than previous average.


##Are there differences in activity patterns between weekdays and weekends?
```{r}
is_weekday <- function (x) {
  !(weekdays(as.Date(x)) %in% c('Sunday', 'Saturday'))
}
data$weekdays <- is_weekday(as.Date(data$date))
steps_perday <- aggregate(data$steps, by= list(interval=data$interval,wd=data$weekdays), FUN=meanrm)

par(mfrow=c(2,1))
plot(steps_perday$x[steps_perday$wd==TRUE], type = 'l', main='Step Pattern in Weekdays')
plot(steps_perday$x[steps_perday$wd==FALSE], type = 'l', main='Step Pattern in Weekend')
```
The patter is apparently different.